\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2016-06-08 paper/ms.tex
Editing and writing about threads
Python model stuff

\section{Python Bindings}\label{sec:python_bindings}

{\bf To be written by Dan FM}

@@

\code{-t <num\_threads>}\\
Run on the specified number of threads. The default is 1.

>>>>

\section{Python Bindings}\label{sec:python_bindings}
In \pkg{DNest4}, it is also possible to specify and run models in
\proglang{Python}. Currently, the only models supported
are ones where the parameters
are a \pkg{NumPy} array of double precision quantities.
Only a single thread is supported, although parallelization at
the level of the \proglang{Python} likelihood function should work.

A Model class implementing the straight line example is given below
and included in {\tt python/examples/straightline/straightline.py}.
The member functions of this class have similar names to those
used in C++, but work slightly differently. Most notably,
an instance of the class does not represent a point in parameter
space, but represents the problem itself. The functions take
\pkg{NumPy} arrays of parameters as input, or produce them as
output, rather than having parameters as data members as in C++.

\begin{CodeChunk}
\begin{CodeInput}
class Model(object):
    """
    Specify the model in Python.
    """
    def __init__(self):
        """
        Parameter values *are not* stored inside the class
        """
        pass

    def from_prior(self):
        """
        Unlike in C++, this must *return* a numpy array of parameters.
        """
        m = 1E3*rng.randn()
        b = 1E3*rng.randn()
        sigma = np.exp(-10.0 + 20.0*rng.rand())
        return np.array([m, b, sigma])

    def perturb(self, params):
        """
        Unlike in C++, this takes a numpy array of parameters as input,
        and modifies it in-place. The return value is still logH.
        """
        logH = 0.0
        which = rng.randint(3)

        if which == 0 or which == 1:
            logH -= -0.5*(params[which]/1E3)**2
            params[which] += 1E3*dnest4.randh()
            logH += -0.5*(params[which]/1E3)**2
        else:
            log_sigma = np.log(params[2])
            log_sigma += 20*dnest4.randh()
            # Note the difference between dnest4.wrap in Python and
            # DNest4::wrap in C++. The former *returns* the wrapped value.
            log_sigma = dnest4.wrap(log_sigma, -10.0, 10.0)
            params[2] = np.exp(log_sigma)

        return logH

    def log_likelihood(self, params):
        """
        Gaussian sampling distribution.
        """
        m, b, sigma = params
        var = sigma**2
        return -0.5*data.shape[0]*np.log(2*np.pi*var)\
                - 0.5*np.sum((data[:,1] - (m*data[:,0] + b))**2)/var

\end{CodeInput}
\end{CodeChunk}

A sampler can then be created and run using the following code.
The \proglang{Python} sampler does not load an {\tt OPTIONS} file.
Instead, the options are provided as function arguments
when setting up the sampler.
\begin{CodeChunk}
\begin{CodeInput}
# Create a model object and a sampler
model = Model()
sampler = dnest4.DNest4Sampler(model,
                               backend=dnest4.backends.CSVBackend(".",
                                                                  sep=" "))

# Set up the sampler. The first argument is max_num_levels
gen = sampler.sample(max_num_levels=30, num_steps=1000, new_level_interval=10000,
                      num_per_step=10000, thread_steps=100,
                      num_particles=5, lam=10, beta=100, seed=1234)

# Do the sampling (one iteration here = one particle save)
for i, sample in enumerate(gen):
    print("# Saved {k} particles.".format(k=(i+1)))
\end{CodeInput}
\end{CodeChunk}

@@

\code{-t <num\_threads>}\\
Run on the specified number of threads. The default is 1.
Be aware that the number of particles specified in the {\tt OPTIONS}
file is actually the number of particles {\em per thread}. Therefore,
if running on 8 or more cores, we recommend setting the number of particles
per thread to 1, so the number of particles equals the number of threads.
